Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2021-12-27 ms/SI.Rmd
Change to equal-area grid cells

```{r map-plot-prep}
# Crop high-res map of Japan to spatial div results area
japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))

# Format biodiv_ferns_spatial data
biodiv_ferns_spatial <-
  biodiv_ferns_spatial %>% 
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super")),
    # can ignore warning about unknown levels, since they don't exist in all variables but that's OK
    across(
      contains("signif"),
      ~fct_relevel(.x, "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
    )
  )

biodiv_ferns_endemic_spatial <-
  biodiv_ferns_endemic_spatial %>% 
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super"))
  )

# Subset ferns biodiversity dataframe: restrict sites to only those used in model
# (drops outlier % apomictic site)
biodiv_ferns_for_plotting <-
  biodiv_ferns_spatial %>%
  drop_apo_outlier

@@

**`r s_figure("envir")`**. Reproductive mode and environmental data in 0.2 degree grid cells across Japan. (a), rolling mean area (km^2^) in 1° latitudinal bands.  (b), % apomictic taxa. (c), annual precipitation. (d), percipitation seasonality (coefficient of variation).  (e), mean annual temperature. (f), temperature seasonality (standard deviation × 100).

@@

```{r raw-lat, fig.width = 7, fig.height = 4}
# Convert spatial data frame to centroids with raw metrics, drop single high apomixis outlier
biodiv_ferns_cent_raw <- spatial_to_cent_for_model(
  biodiv_ferns_spatial,
  c("grids", "lat", "long", "richness", "pd_obs", "fd_obs"))

@@

**`r s_figure("dendrogram")`**. Bioregion dendrograms. Tips are sites (names not shown). Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between sites (see Materials and Methods). Color shows bioregion; bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category. (a), taxonomic bioregions. (b), phylogenetic bioregions. Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].
\clearpage

@@

**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in (a) taxonomic or (b) phylogenetic bioregions arranged by mean annual temperature and annual precipitation. Points shown with partial transparency; darker areas indicate overlapping points. Bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category.

>>>>

```{r map-plot-prep}
# Crop high-res map of Japan to spatial div results area
japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial)) %>%
  st_transform(4612)

# Format biodiv_ferns_spatial data for plotting
biodiv_ferns_spatial <-
  biodiv_ferns_spatial %>% 
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super")),
    # can ignore warning about unknown levels, since they don't exist in all variables but that's OK
    across(
      contains("signif"),
      ~fct_relevel(.x, "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
    )
  ) %>%
  st_transform(4612) %>%
  # Only plot data included in model
  inner_join(select(biodiv_ferns_cent, grids), by = "grids")

biodiv_ferns_endemic_spatial <-
  biodiv_ferns_endemic_spatial %>% 
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super"))
  ) %>%
  st_transform(4612) %>%
  # Only plot data included in model
  inner_join(select(biodiv_ferns_cent, grids), by = "grids")

@@

**`r s_figure("envir")`**. Reproductive mode and environmental data in 20 km × 20 km grid cells across Japan. (a), rolling mean area (km^2^) in 20 km latitudinal bands.  (b), % apomictic taxa. (c), annual precipitation. (d), percipitation seasonality (coefficient of variation). (e), mean annual temperature. (f), temperature seasonality (standard deviation × 100). Data excluded from models not shown in this and subsequent plots unless otherwise mentioned.

@@

```{r raw-lat, fig.width = 7, fig.height = 4}
# Convert spatial data frame to centroids with raw metrics
biodiv_ferns_cent_raw <-
  biodiv_ferns_spatial %>%
  sf_add_centroids() %>%
  st_drop_geometry()

@@

**`r s_figure("dendrogram")`**. Bioregion dendrograms. Tips are sites (names not shown). Dendrograms generated by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between sites (see Materials and Methods). Color shows bioregion; bioregions not consisting of > `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category. (a), taxonomic bioregions. (b), phylogenetic bioregions. Dendrograms visualized with the `r pack("ggtree")` R package [@Yu2017].
\clearpage

@@

**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in (a) taxonomic or (b) phylogenetic bioregions arranged by mean annual temperature and annual precipitation. Points shown with partial transparency; darker areas indicate overlapping points. Bioregions not consisting of > `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.

### 2021-12-27 ms/data_exploration.Rmd
Change to equal-area grid cells

# Inspect the rows with missing data
biodiv_ferns_cent_raw %>% 
  slice(error_rows) %>% 
  select(grids, where(~any(is.na(.x))))
```

The missing data was due to one grid-cell that lacks environmental data. This must be missing from the WorldClim dataset. Let's remove it, then continue.

@@

There are a few points beyond the whiskers that may be outliers. In particular, there is one value of `percent_apo` > 60% that seems quite outside the normal range.

@@

Grubb's test confirms that the percent apomictic value > 60% is an outlier. It also indicates that the highest values of the variables `fd_obs_z`, `pd_obs_z`, `rfd_obs_z`, and `rpd_obs_z` may be outliers.

@@

```{r check-richness-apo}
biodiv_ferns_cent_no_missing %>%
  slice_max(1, order_by = percent_apo) %>%
  select(grids:long, percent_apo, richness)
```

There are just three species in the grid-cell with > 60% apomictic taxa. This data point should be excluded.

@@

They all come from the same two grid-cells, but don't have an extremely low number of species.

>>>>

# Inspect the rows with missing data
biodiv_ferns_cent_raw %>%
  slice(error_rows) %>%
  select(grids, richness, percent_apo, temp, precip, precip_season)
```

The missing data are due to two grid-cells that lack environmental data from the WorldClim dataset and one grid-cell that only had a single species with no breeding-system data. Let's remove these, then continue.

@@

There are a few points beyond the whiskers that may be outliers. In particular, there is one value of `percent_apo` that seems quite outside the normal range.

@@

Grubb's test confirms that the percent apomictic value is an outlier. It also indicates that the highest values of the variables `fd_obs_z`, `pd_obs_z`, `rfd_obs_z`, and `rpd_obs_z` may be outliers.

@@

```{r check-richness-apo}
biodiv_ferns_cent_no_missing %>%
  slice_max(n = 1, order_by = percent_apo, with_ties = TRUE) %>%
  select(grids:long, percent_apo, richness)
```

There is just one species in the grid-cell that obviously contributes to the 100% apomixis rate. This data point should be excluded.

@@

None of these have an extremely low number of species.

### 2021-12-27 ms/data_readme.Rmd
Specify commit hash for targets cache
Update data README date in Rmd
Change to equal-area grid cells

@@

```{r redundancy-by-res-cols, results = "asis"}
  redundancy_by_res %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "res", "Degree of resolution (grid-cell side length), from 0.1 to 0.4 degrees",
      "grids", "Name of grid-cell",
      "abundance", "Number of specimens occurring in grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "redundancy", "Redundancy (measure of sampling completeness)"
  ))
```

@@

```{r japan-ferns-shape-full-cols, results = "asis"}
  japan_ferns_shape_full %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "grids", "Name of grid-cell",
      "abundance", "Number of specimens occurring in grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "redundancy", "Redundancy (measure of sampling completeness)",
      "geom", "Vector describing shape and position of grid-cell"
  ))
```

@@

```{r shape-ferns-dryad-file-cols, results = "asis"}
  shape_ferns_dryad %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "grids", "Name of grid-cell",
      "abundance", "Number of specimens occurring in grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "redundancy", "Redundancy (measure of sampling completeness)",
      "geom", "Vector describing shape and position of grid-cell"
  ))
```

@@

2021-09-09

>>>>

Corresponding commit in repo (https://github.com/joelnitta/japan_ferns_spatial_phy): FIXME add manually

@@

```{r redundancy-by-res-cols, results = "asis"}
  redundancy_by_res %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "res", "Degree of resolution (grid-cell side length), from 0.1 to 0.4 degrees",
      "grids", "Name of grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "abundance", "Number of specimens occurring in grid-cell",
      "redundancy", "Redundancy (measure of sampling completeness)"
  ))
```

@@

```{r japan-ferns-shape-full-cols, results = "asis"}
  japan_ferns_shape_full %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "grids", "Name of grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "abundance", "Number of specimens occurring in grid-cell",
      "redundancy", "Redundancy (measure of sampling completeness)",
      "geom", "Vector describing shape and position of grid-cell"
  ))
```

@@

```{r shape-ferns-dryad-file-cols, results = "asis"}
  shape_ferns_dryad %>%
  make_data_desc(
    tribble(
      ~col, ~desc,
      "grids", "Name of grid-cell",
      "richness", "Number of taxa occurring in grid-cell",
      "abundance", "Number of specimens occurring in grid-cell",
      "redundancy", "Redundancy (measure of sampling completeness)",
      "geom", "Vector describing shape and position of grid-cell"
  ))
```

@@

2021-09-22

### 2021-12-27 ms/manuscript.Rmd
Change to equal-area grid cells

**Methods:** We compiled a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 0.2° grid-cells including `r ncol(comm_ferns)` taxa based on > `r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` specimen records.
We combined this with a phylogeny and functional traits to analyze taxonomic, phylogenetic, and functional diversity, and modeled biodiversity metrics in response to environmental factors and reproductive mode.
We used hierarchical clustering to analyze bioregions.
We compared the overlap of significantly diverse grid-cells with conservation zones and range maps of native Japanese deer to assess conservation status and threats.

@@

repro_count$known$total

@@

We used a list of all native, non-hybrid fern specimens deposited at TNS [herbarium codes follow @Thiers2021]. 
We chose to use this over other sources (e.g., GBIF), because of its high taxonomic quality: all identifications are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/). 
Furthermore, the overall sampling completeness of this dataset is extremely high (`r s_figure("sampling-curve")` in Appendix S1; see Supplemental Data with this article). 
The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa at the rank of species, subspecies, form, or variety (hereafter referred to as "taxa" unless otherwise indicated), excluding non-native taxa and nothotaxa.
Latitude and longitude of each occurrence record were projected into the Japanese Geodetic Datum 2000 (JGD 2000), which was also used for all further geographic analyses.
We filtered out any occurrences not within a map of standard grid squares defined by the Statistics Bureau of Japan (http://www.stat.go.jp/english/data/mesh/05.html) and removed duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).
The R package `r pack("sf")` was used for all GIS analyses [@Pebesma2018].

A necessary step in any analysis of biodiversity is to set the grain size used to accurately define communities. 
Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. 
We tested using 0.1°, 0.2°, 0.3°, and 0.4° grid-cells (size refers to length of side of a grid-cell) for defining communities (co-occurring species). 
At each grain size, species occurrences were converted to a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell). 
We calculated sampling redundancy [1 - richness/number of specimens; @Garcillan2003] to quantify adequacy of sampling. 
Preliminary analysis indicated that 0.2° grid-cells are optimal for our dataset (`r s_figure("grain-size")`). 
We filtered out any cells with redundancy < 0.1 and used this size for subsequent analyses.

@@

We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")` [@R-raster], and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`). 
We identified the intersection of the climate data at the 2.5 minute scale with the 0.2° grid-cells, then calculated the mean of each climatic variable for each 0.2° grid-cell.
We calculated area as rolling mean area (km^2^) across 1° latitudinal bands.
This approach reflects the idea that the amount of area with a given climate regime enclosing each grid-cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by (`r figure("japan-map")`).

@@

# Verify that tips in the plastome tree are accounted for
# by FTOL sampling + japan fern rbcL
tibble(taxon = plastome_tree$tip.label) %>%
  anti_join(select(plastome_sampling, taxon) %>% unique) %>%
  anti_join(tibble(taxon = rownames(japan_rbcL))) %>%
  verify(nrow(.) == 0, success_fun = success_logical)

@@

Initial inspection of the biodiversity and environmental data showed a single outlier in % apomictic taxa due to extremely low species richness (Appendix S2; see Supplemental Data with this article), so that grid-cell and one grid-cell lacking environmental data were excluded from modeling analyses.

@@

```{r dataset-stats}
### Taxon list ###
# Make list of all native ferns in Japan (based off Green List), *including* hybrids
fern_list <-
  green_list %>%
  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
  left_join(ppgi) %>%
  filter(class == "Polypodiopsida")

@@

```{r discus-model-stats}
# Check that conditional AIC was lower in the reproductive model compared to the environmental model for SES of RPD and SES of PD
# Split aic_env_repro dataframe into nested list for easier indexing
c_aic <- aic_env_repro %>% split(.$resp_var) %>% map(~split(., .$model_type))
assert_that(c_aic$pd_obs_z$percent_apo$conditional_aic < c_aic$pd_obs_z$temp$conditional_aic)
assert_that(c_aic$rpd_obs_z$percent_apo$conditional_aic < c_aic$rpd_obs_z$temp$conditional_aic)
```

All models were significant relative to null models without any fixed effects (`r s_table("lrt-null")`). 
The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`A--E), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`F--G); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models. 
Richness showed a hump-shaped relationship with temperature, reaching maximum richness at intermediate temperatures (`r figure("modelfit")`A). 
Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`B--E). 
SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`G); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`F). 
Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD and SES of PD (`r s_table("aic")`).

@@

Clustering analysis grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively. 
The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and northern Honshu), bioregion 2 (southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`A; `r s_figure("dendrogram")`A). 
Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 extended slightly further south (`r figure("bioregions")`B; `r s_figure("dendrogram")`B). 
The other much smaller bioregions ("singletons" including only a single grid-cell each) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.

@@

```{r limits-stats}
# Find the species with the widest latitudinal distribution
taxon_with_widest_lat <- lat_span_summary %>%
  mutate(lat_span = max_lat - min_lat) %>%
  slice_max(lat_span) %>%
  # Add formatted scientific name
  left_join(
    tibble(
      taxon = "Pteridium_aquilinum_japonicum",
      text = "*Pteridium aquilinum* (L.) Kuhn subsp. *japonicum* (Nakai) A. et S. Löve"
    ),
    by = "taxon"
  ) %>%
  # Round off latitude
  mutate(across(c(min_lat, max_lat), ~round(., 2)))

@@

**`r s_table("moran")`**. Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of RPD, SES of FD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.  
**`r s_table("corr")`**. Results of modified *t*-test for correlation between variables while taking into account spatial position.  
**`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan.  
**`r s_table("phy-sig-binary")`**. Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.  
**`r s_table("lrt-null")`**. Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).  
**`r s_table("aic")`**. Model fit as measured with conditional Akaike Information Criterion (cAIC).  
**`r s_figure("sampling-curve")`**. Species collection curve for the ferns of Japan.  
**`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy.    
**`r s_figure("envir")`**. Reproductive mode and environmental data in 0.2° grid cells across Japan.  
**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding nothotaxa.  
**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 0.2° grid cell in the ferns of Japan.  
**`r s_figure("raw-lat")`**. Raw taxonomic diversity, phylogenetic diversity, and functional diversity of the ferns of Japan plotted by latitude.  
**`r s_figure("raw-biplot")`**. Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.  
**`r s_figure("endemism-restricted")`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.  
**`r s_figure("k-plot")`**. Selection of *k* for bioregions.  
**`r s_figure("dendrogram")`**. Bioregion dendrograms.  
**`r s_figure("genus-map")`**. Map of species richness for selected genera of ferns of Japan.  
**`r s_figure("apo-genus")`**. Rate of apomixis in genera of Japanese ferns with > 1 apomictic taxon.  
**`r s_figure("modelfit-not-ult")`**. Relationship between phylogenetic diversity (PD) and relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).  
**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in taxonomic or phylogenetic bioregions arranged by mean annual temperature and annual precipitation.  
**`r s_figure("protected-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and protected areas.  
**`r s_figure("deer-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*).

@@

# Crop high-res map of Japan to spatial div results area
japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial))

# Format biodiv_ferns_spatial data
biodiv_ferns_spatial <-
  biodiv_ferns_spatial %>% 
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(endem_type, levels = c("neo", "paleo", "not significant", "mixed", "super")),
    # can ignore warning about unknown levels, since they don't exist in all variables but that's OK
    across(
      contains("signif"),
      ~fct_relevel(.x, "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
    )
  )
```

@@

**`r figure("bioregions")`**. Bioregions of the ferns of Japan. (A) Taxonomic bioregions. (B) Phylogenetic bioregions. Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid-cells. Bioregions not consisting of > 1 grid-cell each are lumped into the "Other" category.

>>>>

**Methods:** We compiled a community dataset of `r nrow(comm_ferns) %>% scales::number(big.mark = ",")` 20 km × 20 km grid-cells including `r ncol(comm_ferns)` taxa based on > `r occ_sum$filtered$n_specimens$value %>% plyr::round_any(1000, f = floor) %>% number(big.mark = ",")` specimen records.
We combined this with a phylogeny and functional traits to analyze taxonomic, phylogenetic, and functional diversity, and modeled biodiversity metrics in response to environmental factors and reproductive mode.
We used hierarchical clustering to analyze bioregions.
We compared the overlap of significantly diverse grid-cells with conservation zones and range maps of native Japanese deer to assess conservation status and threats.

@@

@@

We used a list of native, non-hybrid fern specimens deposited at herbaria in Japan to quantify occurrences [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
We chose to use this over other sources (e.g., GBIF) because of its extremely high quality, which obviates many cleaning steps otherwise needed when working with large, publicly available datasets.
Furthermore, the overall sampling completeness of this dataset is also extremely high (`r s_figure("sampling-curve")` in Appendix S1; see Supplemental Data with this article). 
The original list comprises `r occ_sum$unfiltered$n_specimens$value %>% scales::number(big.mark = ",")` specimens representing `r occ_sum$unfiltered$n_taxa$value` terminal taxa at the rank of species, subspecies, form, or variety (hereafter referred to as "taxa" unless otherwise indicated), excluding non-native taxa and nothotaxa.
All names are standardized to a common taxonomy, the Japan Green List v. 1.01 (http://www.rdplants.org/gl/).

Occurrence records were georeferenced and thoroughly vetted as follows [@Ebihara2016b; @Ebihara2017; @Ebihara2019b].
For specimens that lacked latitude and logitude data, georeferencing was done by mapping collection site names to a set of standard ca. 10 × 10 km grid squares defined by the Statistics Bureau of Japan (the "second-degree mesh"; http://www.stat.go.jp/english/data/mesh/05.html), and the centroid of the second-degree mesh cell used as the specimen location.
In the case that the collection site could not be mapped to a single second-degree mesh cell, it was excluded.
Next, occurrence maps were generated for each taxon showing presence or absence in each second-degree mesh cell.
In the case that a given taxon appeared insufficiently sampled (i.e., not present on the map where it would typically be expected to occur), AE and members of the Nippon Fernist Club (amateur botanists familiar with the ferns of Japan) searched for additional specimens, which were then added to the list. 
The occurrence maps were iteratively refined until the vast majority of known Japanese ferns had been observed across their expected ranges; the resulting set of occurrence records can be considered accurate to ca. 10 km (the grain size of the second-degree mesh map).

We further cleaned the list prior to analysis by filtering out any occurrences not within the second-degree mesh and removing duplicate collections (`r occ_sum$filtered$n_specimens$value %>% scales::number(big.mark = ",")` specimens, `r occ_sum$filtered$n_taxa$value` taxa after filtering).

A necessary step in any analysis of biodiversity is to set the grain size used to accurately define communities (co-occurring species). 
Here, one must consider that smaller grain size is needed to detect environmental effects, while larger grain size is needed to ensure adequate species sampling. 
We created sets of 10 km, 20 km, 30 km, and 40 km grid-cells covering Japan using a Mollweide equal-area projection (size refers to the length of the side of each square grid-cell).
At each grain size, species occurrences were converted to a presence-absence community matrix (a species was considered present if at least one specimen was recorded in that grid-cell). 
We calculated sampling redundancy [1 - richness/number of specimens; @Garcillan2003] to quantify adequacy of sampling. 
Preliminary analysis indicated that 20 km grid-cells are optimal for our dataset (`r s_figure("grain-size")`). 
Although the grid-cells are defined with equal area, actual land area of each cell varies due to presence of coastlines, etc.
Strictly filtering out all grid-cells with less than complete land area would result in a large loss of data, as Japan has many coastlines and small islands.
Therefore, we instead filtered the grid-cells by sampling completeness, removing any cells with redundancy < 0.1; this dataset was used for all subsequent analyses.
The R package `r pack("sf")` was used for all GIS analyses [@Pebesma2018].

@@

We downloaded climate data at the 2.5 minute scale from the WorldClim database using the `r func("getData")` function in the R package `r pack("raster")` [@R-raster], and extracted four variables related to our research questions as follows: mean annual temperature (bio1; hereafter "temperature"), temperature seasonality (bio4), annual precipitation (bio12; hereafter "precipitation"), and precipitation seasonality (bio15) (`r s_figure("envir")`). 
We identified the intersection of the climate data at the 2.5 minute scale with the 20 km grid-cells, then calculated the mean of each climatic variable for each grid-cell.
We calculated area as rolling mean area (km^2^) across 100 km latitudinal bands.
This approach reflects the idea that the amount of area with a given climate regime enclosing each grid-cell may influence species richness of that cell [@Fahrig2013], and that using raw island areas would not be appropriate in this context as there are just four very large islands with many much smaller ones located close by (`r figure("japan-map")`).

@@

# Verify that tips in the plastome tree are accounted for
# by FTOL sampling + japan fern rbcL
tibble(taxon = plastome_tree$tip.label) %>%
  anti_join(select(plastome_sampling, taxon) %>% unique, by = "taxon") %>%
  anti_join(tibble(taxon = rownames(japan_rbcL)), by = "taxon") %>%
  verify(nrow(.) == 0, success_fun = success_logical)

@@

Two grid cells lacking environmental data and an outlier with 100% apomictic taxa due to extremely low species richness were excluded from modeling analyses (Appendix S2; see Supplemental Data with this article).

@@

```{r dataset-stats}
### Taxon list ###
# Make list of all native ferns in Japan (based off Green List), *including* hybrids
fern_list <-
  green_list %>%
  mutate(genus = str_split(taxon, "_") %>% map_chr(1)) %>%
  left_join(ppgi, by = "genus") %>%
  filter(class == "Polypodiopsida")

@@

```{r discus-model-stats}
# Check that conditional AIC was lower in the reproductive model compared to 
# the environmental model for SES of RPD
# Split aic_env_repro dataframe into nested list for easier indexing
c_aic <- aic_env_repro %>% split(.$resp_var) %>% map(~split(., .$model_type))
assert_that(c_aic$rpd_obs_z$percent_apo$conditional_aic < c_aic$rpd_obs_z$temp$conditional_aic)
# But not for SES of PD
assert_that(c_aic$pd_obs_z$percent_apo$conditional_aic > c_aic$pd_obs_z$temp$conditional_aic)
```

All models were significant relative to null models without any fixed effects (`r s_table("lrt-null")`). 
The effect of temperature dominated (had the greatest absolute *t*-value) in environmental models (`r figure("coeff")`A--E), and % apomictic taxa dominated in reproductive models (`r figure("coeff")`F--G); temperature and % apomictic taxa were significant in each model where they were included, whereas other predictors (area, precipitation, precipitation seasonality) were not significant in some models. 
Richness showed a hump-shaped relationship with temperature, reaching maximum richness at intermediate temperatures (`r figure("modelfit")`A). 
Phylogenetic and functional diversity showed either moderately (SES of PD, SES of FD) to strongly (SES of RPD, SES of RFD) hump-shaped relationships with temperature (`r figure("modelfit")`B--E). 
SES of RPD showed a clear, linear decline with % apomictic taxa (`r figure("modelfit")`G); a similar pattern was detected in SES of PD, but with wider confidence intervals (`r figure("modelfit")`F). 
Conditional AIC was lower (indicating better model fit) in the reproductive model compared to the environmental model for SES of RPD, but not SES of PD (`r s_table("aic")`).

@@

Clustering analysis grouped grid-cells into `r english2(k_taxonomy$optimal$k)` or `r english2(k_phylogeny$optimal$k)` bioregions (`r s_figure("k-plot")`) on the basis of taxonomic or phylogenetic distances, respectively. 
The vast majority of grid-cells fell to a subset of these bioregions: using taxonomic distances, they include bioregion 1 (Hokkaido and northern Honshu), bioregion 2 (southern Honshu, Kyushu, and Shikoku), bioregion 3 (Ryukyu Islands), and bioregion 4 (Ogasawara Islands) (`r figure("bioregions")`A; `r s_figure("dendrogram")`A). 
Phylogenetic distances produced similar results, but the Ryukyu and Ogasawara Islands merged into a single bioregion (3), and bioregion 1 extended slightly further south (`r figure("bioregions")`B; `r s_figure("dendrogram")`B). 
The other much smaller bioregions (including only one or two grid-cells each) are likely artifacts of insufficiently low taxon richness for robust clustering, and are not discussed further.

@@

```{r limits-stats}
# Find the species with the widest latitudinal distribution
taxon_with_widest_lat <- lat_span_summary %>%
  mutate(lat_span = max_lat - min_lat) %>%
  slice_max(lat_span) %>%
  # Add formatted scientific name
  left_join(
    tibble(
      taxon = "Pteridium_aquilinum_japonicum",
      text = "*Pteridium aquilinum* (L.) Kuhn subsp. *japonicum* (Nakai) A. et S. Löve" #nolint
    ),
    by = "taxon"
  ) %>%
  # Round off latitude
  mutate(across(c(min_lat, max_lat), ~round(., 2)))

@@

**`r s_table("moran")`**. Spatial autocorrelation in biodiversity metrics (richness, SES of PD, SES of RPD, SES of FD, SES of RFD) and independent variables (environmental variables and % apomictic taxa) as measured with Moran's *I*.  
**`r s_table("corr")`**. Results of modified *t*-test for correlation between variables while taking into account spatial position.  
**`r s_table("phy-sig")`**. Phylogenetic signal in continuous functional traits of the ferns of Japan.  
**`r s_table("phy-sig-binary")`**. Phylogenetic signal in quantitative (binary) functional traits of the ferns of Japan.  
**`r s_table("lrt-null")`**. Likelihood ratio test (LRT) between full model and null model (model only including the spatial Matérn correlation matrix).  
**`r s_table("aic")`**. Model fit as measured with conditional Akaike Information Criterion (cAIC).  
**`r s_figure("sampling-curve")`**. Species collection curve for the ferns of Japan.  
**`r s_figure("grain-size")`**. Effect of grain size on sampling redundancy.    
**`r s_figure("envir")`**. Reproductive mode and environmental data in 20 km grid cells across Japan.  
**`r s_figure("tree")`**. Time tree of native Japanese ferns and lycophytes, excluding nothotaxa.  
**`r s_figure("abundance")`**. Observed number of specimens and sampling redundancy per 20 km grid cell in the ferns of Japan.  
**`r s_figure("raw-lat")`**. Raw taxonomic diversity, phylogenetic diversity, and functional diversity of the ferns of Japan plotted by latitude.  
**`r s_figure("raw-biplot")`**. Relationships between observed functional and phylogenetic diversity and taxonomic richness in the ferns of Japan.  
**`r s_figure("endemism-restricted")`**. Phylogenetic endemism of the ferns of Japan measured using CANAPE (categorical analysis of neo- and paleo-endemism), restricted dataset including only taxa endemic to Japan.  
**`r s_figure("k-plot")`**. Selection of *k* for bioregions.  
**`r s_figure("dendrogram")`**. Bioregion dendrograms.  
**`r s_figure("genus-map")`**. Map of species richness for selected genera of ferns of Japan.  
**`r s_figure("apo-genus")`**. Rate of apomixis in genera of Japanese ferns with > 1 apomictic taxon.  
**`r s_figure("modelfit-not-ult")`**. Relationship between phylogenetic diversity (PD) and relative PD and % apomictic taxa, models inferred using phylogeny with branchlengths in units of expected genetic change (untransformed ML tree).  
**`r s_figure("bioregions-env")`**. Scatterplots of grid cell membership in taxonomic or phylogenetic bioregions arranged by mean annual temperature and annual precipitation.  
**`r s_figure("protected-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and protected areas.  
**`r s_figure("deer-map")`**. Maps showing overlap of grid-cells with significantly high biodiversity for the ferns of Japan and distribution of Japanese deer (*Cervus nippon*).

@@

# Crop high-res map of Japan to spatial div results area
japan <- sf::st_crop(japan_shp, sf::st_bbox(biodiv_ferns_spatial)) %>%
  st_transform(4612)

# Format biodiv_ferns_spatial data for plotting
biodiv_ferns_spatial <-
  biodiv_ferns_spatial %>%
  # Set order of factor levels for plotting
  mutate(
    endem_type = factor(
      endem_type,
      levels = c("neo", "paleo", "not significant", "mixed", "super")),
    # can ignore warning about unknown levels, since they don't exist in all
    # variables but that's OK
    across(
      contains("signif"),
      ~fct_relevel(.x, 
      "< 0.01",  "< 0.025", "not significant", "> 0.975", "> 0.99")
    )
  ) %>%
  st_transform(4612)
```

@@

**`r figure("bioregions")`**. Bioregions of the ferns of Japan. (A) Taxonomic bioregions. (B) Phylogenetic bioregions. Bioregions determined by clustering taxonomic (Sørensen) or phylogenetic (PhyloSor) distances between grid-cells. Bioregions not consisting of > `r english2(bioregion_cutoff)` grid-cells each are lumped into the "Other" category.
